\documentclass[12pt, letterpaper]{article}

\usepackage[letterpaper,left=0.7in,right=0.7in,top=\dimexpr15mm+1.5\baselineskip,bottom=0.7in]{geometry}
\usepackage{amsmath}
\usepackage{pxfonts}
\usepackage[T1]{fontenc}

\linespread{1.5}

\begin{document}

$$ \int \arctan{t} \, dt $$

Let $u = \arctan{t}$, then clearly (make sure you see that it \emph{is} clear) $dv = dt.$ Then

\begin{equation*}
\begin{aligned}
u = \arctan{t} \quad &\Rightarrow\quad du = \frac{dt}{1 + t^2} \\
dv = dt \quad &\Rightarrow\quad v = t
\end{aligned}
\end{equation*}

so that

\begin{equation*}
\begin{aligned}
\int \arctan{t} \, dt
  &= uv - \int v \, du \\[0.2in]
  &= t \arctan{t} - \int \frac{t \, dt}{1 + t^2}
\end{aligned}
\end{equation*}

Now use $u$-substitution with

$$ u = 1 + t^2 \quad \Rightarrow \quad \tfrac{1}{2}du = t \, dt $$

which gives

\begin{equation*}
\begin{aligned}
\int \arctan{t} \, dt
  &= t \arctan{t} - \int \frac{t \, dt}{1 + t^2} \\[0.2in]
  &= t \arctan{t} - \tfrac{1}{2} \int \frac{du}{u} \\[0.2in]
  &= t \arctan{t} - \tfrac{1}{2} \log{|u|} + C \\[0.2in]
  &= t \arctan{t} - \tfrac{1}{2} \log{|1 + t^2|} + C.
\end{aligned}
\end{equation*}

So we have (Where did the square root come from? Where did the absolute value go?)

\begin{equation*}
\boxed{
\int \arctan{t} \, dt
  = t \arctan{t} - \log{\sqrt{1 + t^2}} + C.
}
\end{equation*}

\newpage

$$ \int \cos{\left(\log{t}\right)} \, dt $$

The key observation here is that, since $\frac{d \log{t}}{dt} = \frac{1}{t}$ and $\int dt = t$, those two terms will cancel out when we do our integration by parts (take a second to try to see it). You would have figured this out by experimentation eventually, but better if you could see it before you got started. So now let

\begin{equation*}
\begin{aligned}
u = \cos{\left(\log{t}\right)} \quad &\Rightarrow\quad du = -\tfrac{1}{t} \sin{\left( \log{t} \right)} \, dt \\
dv = dt \quad &\Rightarrow\quad v = t
\end{aligned}
\end{equation*}

so that

\begin{equation*}
\begin{aligned}
\int \cos{\left(\log{t}\right)} \, dt
  &= uv - \int v \, du \\[0.2in]
  &= t\cos{\left(\log{t}\right)} - \int t \left( -\tfrac{1}{t} \sin{\left( \log{t} \right)} \right) \, dt \\[0.2in]
  &= t\cos{\left(\log{t}\right)} + \int \sin{\left( \log{t} \right)} \, dt \\[0.2in]
\end{aligned}
\end{equation*}

Now do another integration by parts with

\begin{equation*}
\begin{aligned}
u = \sin{\left(\log{t}\right)} \quad &\Rightarrow\quad du = \tfrac{1}{t} \cos{\left( \log{t} \right)} \, dt \\
dv = dt \quad &\Rightarrow\quad v = t
\end{aligned}
\end{equation*}

to get

\begin{equation*}
\begin{aligned}
\int \cos{\left(\log{t}\right)} \, dt
  &= t\cos{\left(\log{t}\right)} + \int \sin{\left( \log{t} \right)} \, dt \\[0.2in]
  &= t\cos{\left(\log{t}\right)} + t\sin{\left(\log{t}\right)} - \int \cos{\left( \log{t} \right)} \, dt.
\end{aligned}
\end{equation*}

Rearranging gives

\begin{equation*}
\begin{aligned}
2\int \cos{\left(\log{t}\right)} \, dt
  = t\cos{\left(\log{t}\right)} + t\sin{\left(\log{t}\right)}
\end{aligned}
\end{equation*}

so that

\begin{equation*}
\boxed{
\int \cos{\left(\log{t}\right)} \, dt
  = \frac{t\cos{\left(\log{t}\right)} + t\sin{\left(\log{t}\right)}}{2}.
}
\end{equation*}

\newpage

$$ \int t\sin{t}\cos{t} \, dt $$

This one wouldn't be bad if we could drop the $t$ term in the front. It would "disappear" (by turning into a $1$) if we took it's derivative, so let's make that our $u$ in our integration by parts:
\begin{equation*}
\begin{aligned}
u = t \quad &\Rightarrow \quad du = dt \\
dv = \sin{t}\cos{t} \, dt \quad &\Rightarrow \quad ???
\end{aligned}
\end{equation*}

Now it remains to integrate our $dv$ term. We can do this with $u$-substitution, but we just used $u$ so lets call it $w$-substitution instead to minimize confusion. The derivative of $\sin{t}$ is $\cos{t}$ so let

$$ w = \sin{t} \quad \Rightarrow \quad dw = \cos{t} \, dt $$

Then we have

\begin{equation*}
\begin{aligned}
v &= \int dv = \int \sin{t}\cos{t} \, dt \\[0.2in]
  &= \int w \, dw = \frac{w^2}{2} \\[0.2in]
  &= \tfrac{1}{2} \sin^2{t}.
\end{aligned}
\end{equation*}

Putting it together, our original integration by parts gives

\begin{equation*}
\begin{aligned}
\int t\sin{t}\cos{t} \, dt
  &= uv - \int v \, du \\[0.2in]
  &= \frac{t}{2} \sin^2{t} - \frac{1}{2} \int \sin{t} \, dt \\[0.2in]
  &= \frac{t}{2} \sin^2{t} + \frac{1}{2} \cos{t} + C.
\end{aligned}
\end{equation*}

Our solution is

\begin{equation*}
\boxed{
\int t\sin{t}\cos{t} \, dt
  = \frac{t\sin^2{t} + \cos{t}}{2} + C.
}
\end{equation*}

\newpage

$$ \int \arctan{\frac{1}{t}} \, dt $$

We don't know the integral of this thing, but we \emph{do} know it's derivative, so to use IBP, it would make sense to let $u = \arctan{\frac{1}{t}}$. In this case, looking at our $dv$ term first we get

$$ dv = dt \quad \Rightarrow \quad v = t$$

Since $u = \arctan{\frac{1}{t}}$, by chain rule we get

\begin{equation*}
\begin{aligned}
du &= -\frac{1}{t^2} \left( \frac{1}{1 + \frac{1}{t^2}} \right) \, dt
    = -\frac{1}{t^2} \left( \frac{t^2}{1 + t^2} \right) \, dt
    = - \frac{dt}{1+t^2}.
\end{aligned}
\end{equation*}

Now our IBP gives

\begin{equation*}
\begin{aligned}
\int \arctan{\frac{1}{t}} \, dt
  &= uv - \int v \, du \\[0.2in]
  &= t \arctan{\frac{1}{t}} - \int t \left( \frac{-dt}{1+t^2} \right) \\[0.2in]
  &= t \arctan{\frac{1}{t}} + \int \frac{t \, dt}{1 + t^2}
\end{aligned}
\end{equation*}

Using $u$-substitution on this last integral we let

$$ u = 1 + t^2 \quad \Rightarrow \quad \frac{du}{2} = t \, dt$$

which gives

\begin{equation*}
\begin{aligned}
\int \arctan{\frac{1}{t}} \, dt
  &= t \arctan{\frac{1}{t}} + \int \frac{t \, dt}{1 + t^2} \\[0.2in]
  &= t \arctan{\frac{1}{t}} + \frac{1}{2} \int \frac{du}{u} \\[0.2in]
  &= t \arctan{\frac{1}{t}} + \frac{1}{2} \log{|1 + t^2|} + C. \\[0.2in]
\end{aligned}
\end{equation*}

Finally we have

\begin{equation*}
\boxed{
\int \arctan{\frac{1}{t}} \, dt
  = t \arctan{\frac{1}{t}} + \log{\sqrt{1 + t^2}} + C.
}
\end{equation*}

\newpage

$$ \int 3^{t}\sin{t} \, dt $$

If you've done a few of these, you'll see that this is a "circular" IBP problem where we will end up with a term that is a factor of the original intergral--similar the ones where we get a "$2I$" in class then divide by $2$ to get our final answer. Terms like $3^t$ work like $e^t$ (\emph{i.e.}, they don't "go away" by integration or differentiation) but with a messy coefficient. So we will use that term as our $dv$.

\begin{equation*}
\begin{aligned}
u = \sin{t} \quad &\Rightarrow \quad du = \cos{t} \, dt \\
dv = 3^t \, dt \quad &\Rightarrow \quad v = \tfrac{1}{\log{3}} \, 3^t
\end{aligned}
\end{equation*}

Where $\tfrac{1}{\log{3}}$ is just a number (it's about $0.91$, and you should double check that on your calculator). So now we have

\begin{equation*}
\begin{aligned}
\int 3^{t}\sin{t} \, dt
  &= \tfrac{1}{\log{3}} \, 3^t \sin{t} - \tfrac{1}{\log{3}} \int 3^t \cos{t} \, dt
\end{aligned}
\end{equation*}

Now we do another IBP on this right-most integral with

\begin{equation*}
\begin{aligned}
u = \cos{t} \quad &\Rightarrow \quad -du = \sin{t} \, dt \\
dv = 3^t \, dt \quad &\Rightarrow \quad v = \tfrac{1}{\log{3}} \, 3^t
\end{aligned}
\end{equation*}

which gives

\begin{equation*}
\begin{aligned}
\int 3^{t}\sin{t} \, dt
  &= \tfrac{1}{\log{3}} \, 3^t \sin{t} - \tfrac{1}{\log{3}} \int 3^t \cos{t} \, dt \\[0.2in]
  &= \tfrac{1}{\log{3}} \, 3^t \sin{t} - \tfrac{1}{(\log{3})^2} \, 3^t \cos{t} - \tfrac{1}{(\log{3})^2} \int 3^{t}\sin{t} \, dt
\end{aligned}
\end{equation*}

Rearranging terms gives

\begin{equation*}
\begin{aligned}
\left( 1 + \tfrac{1}{(\log{3})^2} \right) \int 3^{t}\sin{t} \, dt
  = \tfrac{1}{\log{3}} \, 3^t \sin{t} - \tfrac{1}{(\log{3})^2} \, 3^t \cos{t}
\end{aligned}
\end{equation*}

Finally,

\begin{equation*}
\boxed{
\int 3^{t}\sin{t} \, dt
  = \tfrac{\log{3}}{1 + (\log{3})^2} \left( \sin{t} - \tfrac{1}{\log{3}} \cos{t} \right) 3^t
  \approx 0.5 \left( \sin{t} - 0.91 \cos{t} \right)  3^t  .
}
\end{equation*}

\emph{Whew!}

\end{document}




